---
title: "Homework Week 1 Day 2"
output: html_notebook
---


```{r}
library(tidyverse)
```

```{r}
tidyverse_logo()
```
```{r}
tidyverse::tidyverse_logo()
```

```{r}
devtools::install_github("codeclan/CodeClanData")
```

```{r}
devtools::install_github("codeclan/CodeClanData")
```

```{r}
class(CodeClanData::starwars)
```

```{r}
install.packages("devtools")
```

```{r}
devtools::install_github("codeclan/CodeClanData")
```

```{r}
library(CodeClanData)
```

```{r}
starbucks_data <- drinks_content
```

# Question 1

```{r}
dim(drinks_content)

# Gives the dimensions of the data
```

```{r}
names(drinks_content)

# Gives the variable names
```

```{r}
str(drinks_content)

#Down at the bottom the variable types are shown
```

# Question 2

```{r}
new_table <- select(starbucks_data, Beverage_category, Beverage, Beverage_prep, Calories)

new_table
  
```
# Answer to Question 2

```{r}
starbucks_data %>%
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>%
  filter(is.na(Calories)) %>%
  summarise(count_of_missing_calories = n())
```


# Question 3 

```{r}

new_table_1 <- mutate(new_table, Calorie_Diff = Calories - 135)

new_table_1
```

# Question 4 Summarise the mean number of calories in each beverage_category. Which 3 category of drinks have the most calories? Which 3 drinks have the least? Write a small summary of your findings.

```{r}
new_table_1 %>%
  group_by(Beverage_category) %>%
  summarise(mean_calories = mean(Calories)) %>%
  arrange(desc(mean_calories)) %>%
  head(3)
  

# Can write arrange(desc(`mean(Calories)`) 
```

# Question 5 Let’s look at this a different way. What is the average number of calories in each Beverage_prep type?

```{r}
new_table_1 %>%
  group_by(Beverage_prep) %>%
  summarise(mean)
```

# Question 6: Which Beverage_prep type contains more than the average calories of all drinks?

```{r}
avg_cal <- starbucks_data %>%
  summarise(mean_cal = mean(Calories)) %>%
  pull()


class(avg_cal)
```

```{r}

```

# Question 7:

```{r}
starbucks_data %>%
  filter(Beverage_category != "Tazo® Tea Drinks") %>%
  group_by(Beverage_category, Beverage, Beverage_prep) %>%
  summarise(min_cal = min(Calories)) %>%
  arrange(min_cal)
  head(10)

```

