---
title: "R Notebook"
output: html_notebook
---

```{r}
country_data <- read_csv("data/country_data.csv")
```
```{r}
country_data
```

# Question 1: Load in the country_data.csv file and take a look at it. What needs changed in this to make it tidy? Which columns need made into long format, and which need spread to rows?

# - cases and population needed to be moved up into columns

# - years need to brought down into variables, years. 

# Split up country to country and ID

# Question 2: Do the appropriate pivoting of your columns using R

```{r}
country_data_edited <- country_data %>%
  
  pivot_longer(cols = starts_with("Y"),
               names_to = "year",
               values_to = "number_count",
               names_prefix = "Y",
               names_transform = list(year = as.numeric)) %>%
  
# mutate(year = as.numeric(year)) %>%

# want to change from character to dbl for years
  
# mutate() function
  
# can use name_transform within the pivot_longer
  
  pivot_wider(names_from = type,
              values_from = number_count)

country_data_edited
```

# Question 3: Separate the country variable so you have the name in one column and the number in another.

```{r}
country_data_edited %>%
  
  separate(country,
           c("country_number", "country_name"),
           sep = "\\. ") %>%
  
  unite(code, c("code_letter", "code_numb"),
        sep = "")
```

