---
title: "modelling_practical"
author: "Jonny Nelson"
date: "17/03/2022"
output: html_document
---

# Disconnecting from Spark Previous

```{r}
# disconnect our individual connection
# spark_disconnect(sc)

# or disconnect all open connections
# spark_disconnect_all()
```

## Connecting the Spark and loading sparklyr

```{r}
library(sparklyr)
sc <- spark_connect(master = "local")
```

## Loading the other libraries required

```{r}
library(tidyverse)
```

## Loading the data into Spark

```{r}
profiles <- spark_read_csv(
  sc, 
  "data/profiles.csv",
  escape = "\"",
  memory = FALSE,
  options = list(multiline = TRUE)
)

glimpse(profiles)
```

* memory = FALSE, options = list(multiline = TRUE): To split the essay into 10 fields essay0 - essay9

## Checking for NA values

```{r}
profiles %>%
  summarise_all(.funs = ~ sum(as.integer(is.na(.)))) 
```

* We need to use as.integer() here, as we need to type-convert logical to integer as Spark is a statistical language.

* A lot of NA's here

### Leaving out the as.integer()

```{r}
profiles %>%
  summarise_all(.funs = ~ sum(is.na(.)))
```

* Error as expected 

```{r}
profiles %>%
  summarise(num_neg_income = sum(as.integer(income < 0)))
```

* A decent number of negative incomes as well

# Split into Character and Numeric table for cleaning

## Character Variable Table

```{r}
profiles_char <- profiles %>%
  select(-c(age, income, height)) %>%
  mutate_all(~ ifelse(is.na(.), "missing", .))
```

## Numeric variable Table

```{r}
profiles_num <- profiles %>%
  select(age, income, height) %>%
  mutate(
    age = as.numeric(age),
    income = ifelse(income == "-1", NA, as.numeric(income)),
    height = as.numeric(height)
  )
```

## sdf_bind Back Together

```{r}
profiles <- sdf_bind_cols(profiles_char, profiles_num) %>%
  compute("profiles") %>%
  glimpse()

# The two tables split up no longer have the same number of rows!!! need to trouble shoot.
```

**The two tables split up no longer have the same number of rows!!! need to trouble shoot.**

**At this point continuing on with the practical without running the code**

## Another look at the NA values

```{r}
profiles %>%
  summarise_all(~sum(as.integer(is.na(.))))
```

## Another look at negative income values

```{r}
profiles %>%
  summarise(num_neg_income = sum(as.integer(income < 0)))
```

## Check of the profile schema 

```{r}
schema <- sdf_schema(profiles)

schema <- schema %>%
  transpose() %>%
  as_tibble() %>%
  unnest(cols = c(name, type))

schema
```

## Exploring the categoricals: cross-tabulation

```{r}
contingency <- profiles %>%
  sdf_crosstab("drinks", "drugs") %>%
  collect()

contingency
```

## Reording the Contingency Table above using fct_relevel()

```{r}
contingency %>%
  rename(drinks = drinks_drugs) %>%
  mutate(
    drinks = as_factor(drinks) %>% 
      fct_relevel("missing", "not at all", "rarely", "socially", "often",
                  "very often", "desperately")
  ) %>% 
  arrange(drinks) %>%
  select(drinks, missing, never, sometimes, often)
```

## Mutating the Response Variable in

```{r}
profiles <- profiles %>%
  mutate(
    not_working = as.integer(ifelse(job %in% c("student", "unemployed", "retired"), 1, 0))
  )

profiles %>%
  count(not_working)
```

# Feature Engineering

## Religion: Extracting words

## Too many distinct observations to make Dummy Variables

```{r}
profiles %>%
  distinct(religion)
```

## Extracting the first word out the character string

```{r}
profiles <- profiles %>%
  mutate(religion = regexp_extract(religion, "[a-zA-Z]+", 0)) %>%
  compute("profiles") %>%
  glimpse()
```

## Does prop_not_working vary with religion

```{r}
profiles %>%
  count(religion, not_working) %>%
  group_by(religion) %>%
  summarise(
    count = sum(n),
    prop_not_working = sum(not_working * n) / sum(n)
  )
```

## Essays: Extracting total length

```{r}
profiles <- profiles %>%
  mutate(
    essay_length = char_length(paste(!!!syms(paste0("essay", 0:9))))
  ) %>% 
  compute("profiles") %>%
  glimpse()
```

* syms() function with !!! operator to do the above.
* Gives the character length of all essays put back together

## Converting categoricals to dummies

```{r}
profiles <- profiles %>%
    ft_string_indexer(
    input_col = "drinks",
    output_col = "drinks_indexed"
  ) %>%
  ft_one_hot_encoder(
    input_cols = "drinks_indexed",
    output_cols = "drinks_encoded"
  ) %>%
  ft_string_indexer(
    input_col = "drugs",
    output_col = "drugs_indexed"
  ) %>%
  ft_one_hot_encoder(
    input_cols = "drugs_indexed",
    output_cols = "drugs_encoded"
  ) %>%
  ft_string_indexer(
    input_col = "status",
    output_col = "status_indexed"
  ) %>%
  ft_one_hot_encoder(
    input_cols = "status_indexed",
    output_cols = "status_encoded"
  ) %>%
  compute("profiles")
```

