---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
actors <- read_csv("data/actors.csv")

movies <- read_csv("data/movies.csv")

roles <- read_csv("data/roles.csv")
```

Question 2.

What are the primary keys and the foreign keys for each data frame?
What kind of relationship (one-to-one, one-to-many, many-to-many) is there between
actors and roles
movies and roles
actors and movies

```{r}
# Across actors and movies, "id" is the primary key and then for roles "actor_id" is the foreign key.

# actors and roles - one to many

# movies and roles - one to many

# actors and movies - many to many
```


# Part 2

#Question 1.
#From the movies table, find the id of the film The Little Mermaid. (In the data as “Little Mermaid, The”).

```{r}

# Use the filter function when you open the data frame

```

#Question 2.
#Filter the roles data frame, so we just have roles from the Little Mermaid (hint: use the ID you found in the last question). Save the result as little_mermaid_roles.

```{r}

mermaid_roles <- roles %>%
  filter(movie_id == 192017)
  
```

#Question 3.
#Carry out a left join on this new data frame and the actors data frame, to find the actors which appeared in The Little Mermaid.

```{r}
left_join(mermaid_roles, actors, by = c("actor_id" = "id"))
```

#Question 4.
#Do the last join, but with an inner join instead. Is the data different? Why is that?

```{r}
inner_join(mermaid_roles, actors, by = c("actor_id" = "id"))

# No missing information, so extra observations.
```

# Question 1.
# For all actors who have appeared in at least two films. Call this data frame actors_more_than_once.

```{r}

```


# Question 2.
# Left join actors_more_than_once to roles, so we can see all the actors roles those actors played in.



# Question 3.
# Left join the result from above to movies, so we can see details of each movie for these roles

# Question 4.
# For each actor, find the average rating of the films they appeared. Find the five actors with the highest average rank of films they appeared in.

4