---
title: "Data Exploration"
author: "Jonny Nelson"
date: "12/01/2022"
output: html_document
---

# Task 

## gain better understanding of the incidence of cancer in NHS Borders
## help inform the planning for provision of cancer treatment <br>

# Plan

## * Load in relevant data sets from PHS
## * Filter for NHS Borders using geographical key
## * Better understand what is going on which incidences of cancer - explore the driving factors

# Load in Libraries

```{r}
library(janitor)
library(tidyverse)
library(lubridate)
library(cowplot)
```

# Load in the data sets

```{r}
incidence_regions <- read_csv("raw_data/incidence_by_health_board.csv")

incidence_scotland <- read_csv("raw_data/incidence_at_scotland_level.csv")

clean_scotland <- incidence_scotland %>%
  clean_names()

head(incidence_regions)

glimpse(incidence_regions)

unique(incidence_regions$HB)
# NHS Borders -  . . . 16

unique(incidence_regions$CancerSiteICD10Code)

unique(incidence_regions$CancerSite)

unique(incidence_regions$SexQF)

incidence_regions %>%
  select(CancerSiteICD10Code, CancerSite) %>%
  filter(CancerSiteICD10Code == "C44")

# Therefore C44 - Non-melanoma skin cancer no included in all cancers
# Why?

unique(incidence_regions$WASRLower95pcConfidenceIntervalQF)
# NA and z, z meaning that the number is not statistically significant 

incidence_regions %>%
  filter(WASRLower95pcConfidenceIntervalQF == "z")

# Where z is found there is no confidence intervals

# appears to be true for rates that are very low

# same for europe and world data
```

# HB - Health Board
# CancerSiteIDC10 - diagnosed cancer site
# CancerSite - Type of cancer (aggregated data here watch out)
# Sex - aggreated All then split to male and female (filter out all)
# SexQF - d = derived from other data (can filter SexQF == d as this removes aggregated by sex data)
# Year - Year of diagnosis - mutate to year(year)
# IncidencesAllAges - Total number of new cancer registrations

# Crude rate - per 100,000 
# Crude rate Upper and Lower - CI intervals 

# EASR - European age-standardised (2013 ES Population) per 100,000

# WASR - World age-standardised incidence per 100,000

# SIR - Standardised incidence ratio 


# Initial Cleaning 

```{r}
# Initial cleaning of the data set to mutate in a new variable to show the health boards by name and the clean the names of the variables into the tidy data format.

# Written this clean data into a separate folder and recalled at a later stage

incidence_clean <- incidence_regions %>%
  clean_names() %>%
    mutate(health_board_name = case_when(
    hb == "S92000003" ~ "Scotland",
    hb == "S08000015" ~ "Ayrshire and Arran",
    hb == "S08000016" ~ "Borders",
    hb == "S08000017" ~ "Dumfries and Galloway",
    hb == "S08000019" ~ "Forth Valley",
    hb == "S08000020" ~ "Grampian",
    hb == "S08000022" ~ "Highland",
    hb == "S08000024" ~ "Lothian",
    hb == "S08000025" ~ "Orkney",
    hb == "S08000026" ~ "Shetland",
    hb == "S08000028" ~ "Western Isles",
    hb == "S08000029" ~ "Fife",
    hb == "S08000030" ~ "Tayside",
    hb == "S08000031" ~ "Greater Glasgow and Clyde",
    hb == "S08000032" ~ "Lanarkshire",
    TRUE ~ as.character(NA)))

write_csv(incidence_clean, "clean_data/incidence_clean")
```

# Initial Data Exploration:

## Major Groupings of Cancer Type

```{r}
# Firstly wanted to take a look at just the Borders in a time series idenifying the trend in the crude rate of cancer per 100,000.

# As we can see the trend clearly indicates an increase over time for All cancer types from 1995 to 2019

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All" & cancer_site == "All cancer types") %>%
  ggplot() +
  aes(x = year, y = crude_rate) +
  geom_point() +
  geom_line() +
  geom_ribbon(aes(ymin = crude_rate_lower95pc_confidence_interval,
                  ymax = crude_rate_upper95pc_confidence_interval),
              linetype = 2,
              alpha = 0.1,
              colour = "red",
              fill = "red")
  

# Find cancer types and health board name

# unique(incidence_clean$cancer_site)

# unique(incidence_clean$health_board_name)

# Junk Code

# incidence_clean %>%
#   filter(cancer_site == c("All cancer types", "Bladder", "Bone and articular cartilage", "Bone and connective tissue", "Connective tissue", "Malignant brain cancer", "Malig brain ca (incl pit. gland, cranio. duct, pineal gland)", "Non-malig brain ca (incl pit.gland,cranio.duct,pineal gland)", "All brain and CNS tumours (malignant and non-malignant)") &
#          health_board_name == "Borders") %>%
#   arrange(cancer_site)
```

# Which Types of Cancer are driving this trend from 550 -> 800?

```{r}
# Next I wanted to identify the main drivers in this overall trend increase to get a better understanding of which specific types of cancer were driving the steady increase

# Plotting this out on a using the facet wrap gave too many graphs to be able to discern any trends. So I had to look about things in another way. Afterwards i figured out the solution was to make the plot into a function and then pass the function through plot_grid where you can select the plots shown.

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All" & cancer_site != "All cancer types") %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free") 

  # plot_grid(plotlist = [10:15], nrow = 3, ncol = 3)

# Cannot discern anything from this data so needs to be broken down to a greater degree
  
```

# What cancers had the highest change between 1995 and 2020?

```{r, message = FALSE}
# I was able to calculate the absolute change and percentage change between 1995 and 2019, by using a summarise function. I then arranged the data by percent change and chose the top 20 values in terms of percent change to take a look at closer.

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>%
  group_by(cancer_site) %>%
  summarise(percent_change_in_incidences = (incidences_all_ages/lag(incidences_all_ages, n = 24) - 1) *100,
            absolute_change_in_incidences = diff(incidences_all_ages, lag = 24) ) %>%
  arrange(desc(percent_change_in_incidences)) %>%
  head(21)

# diff(incidences_all_ages, lag = 24) 

# Take a look at percent_change_in_incidences == "lnf"

# * All brain and CNS = 0 -> absolute value
# * Non-malignant brain cancer = 0 -> 10  
# * Salivary glands = 0 -> 2 - still rare

# incidence_clean %>%
#   filter(health_board_name == "Borders") %>%
#   filter(sex == "All") %>%
#   filter(cancer_site == "Salivary glands")
  
```

# Choosing out facets to take a look at (1,2)

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("All brain and CNS tumours (malignant and non-malignant)", "Non-malig brain ca (incl pit.gland,cranio.duct,pineal gland)")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free") 
```

# Top 3,4

```{r}

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Salivary glands", "Carcinoma in situ of the breast")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free") 

```

# 5,6

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Tongue", "Connective tissue")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free") 
```
# 7,8

```{r}
# "Liver and intrahepatic bile ducts"

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Thyroid", "Liver and intrahepatic bile ducts")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")

# Liver is trending up
```
# 9,10

```{r}
# "Multiple myeloma and malignant plasma cell neoplasms", "Oropharyngeal cancers"

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Multiple myeloma and malignant plasma cell neoplasms", "Oropharyngeal cancers")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

# 11,12

```{r}
# Kidney

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Kidney", "Bone and connective tissue")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

# 13,14

```{r}
# Basal cell carcinoma of the skin

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Chronic lymphocytic leukaemia", "Basal cell carcinoma of the skin")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

# 15,16

```{r}
# "Non-melanoma skin cancer", "Pancreas"

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Non-melanoma skin cancer", "Pancreas")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

# 17,18

```{r}
# Lip, oral cavity and pharynx

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Breast", "Lip, oral cavity and pharynx")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

# 19, 20

```{r}
# Squamous cell carcinoma of the skin

incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Squamous cell carcinoma of the skin", "Acute lymphoblastic leukaemia")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

# From all 20 graphs I was able to pick out 9 cancer sites which showed consistent trend increases in incidences in the Borders. Moving forward, it would be preferable to do a linear regression to obtain the gradient, which would inform the expected rate increase that NHS Borders could expect year on year - for each group of cancer

# This could be very helpful to supply chain teams within hospitals who need to provision the treatments for these types of cancer.

# It would also inform hospital management and specialists on how they would be able to manage say, increased chemotherapy, for these expected patients.

# This could inform NHS Borders public health policy and messaging. There may be increased cancer awareness and early detection targeting these cancer types - informed by specialists and consultants - with what to look out for.

# Future - could use the age range statistics to inform what ages and sexs need targeting for messaging


# Breakdowns of Borders vs. Scotland vs. Europe

