---
title: "PHS Scotland Task"
author: "Jonny Nelson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
```

# Task 

<br>

#### _Task_ - Gain better understanding of the incidence of cancer in NHS Borders
#### _Outcome_ - Help inform the planning for provision of cancer treatment 

<br>

```{r, message = FALSE, warning = FALSE}
library(janitor)
library(tidyverse)
library(lubridate)
library(cowplot)

incidence_regions <- read_csv("raw_data/incidence_by_health_board.csv")

incidence_scotland <- read_csv("raw_data/incidence_at_scotland_level.csv")

incidence_clean <- incidence_regions %>%
  clean_names() %>%
    mutate(health_board_name = case_when(
    hb == "S92000003" ~ "Scotland",
    hb == "S08000015" ~ "Ayrshire and Arran",
    hb == "S08000016" ~ "Borders",
    hb == "S08000017" ~ "Dumfries and Galloway",
    hb == "S08000019" ~ "Forth Valley",
    hb == "S08000020" ~ "Grampian",
    hb == "S08000022" ~ "Highland",
    hb == "S08000024" ~ "Lothian",
    hb == "S08000025" ~ "Orkney",
    hb == "S08000026" ~ "Shetland",
    hb == "S08000028" ~ "Western Isles",
    hb == "S08000029" ~ "Fife",
    hb == "S08000030" ~ "Tayside",
    hb == "S08000031" ~ "Greater Glasgow and Clyde",
    hb == "S08000032" ~ "Lanarkshire",
    TRUE ~ as.character(NA)))
```

# Overall Trend in Borders Crude Rate per 100,000
<br>

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All" & cancer_site == "All cancer types") %>%
  ggplot() +
  aes(x = year, y = crude_rate) +
  geom_point() +
  geom_line() +
  geom_ribbon(aes(ymin = crude_rate_lower95pc_confidence_interval,
                  ymax = crude_rate_upper95pc_confidence_interval),
              linetype = 2,
              alpha = 0.1,
              colour = "red",
              fill = "red") +
  ylab("Crude Rate") +
  xlab("Year") +
  ggtitle("Scottish Borders Crude Rate of Cancer Incidences per 100,000")
```
<br>

#### - Initial cleaning of the data set to mutate in a new variable to show the health boards by name and the clean the names of the variables into the tidy data format.

#### - Firstly wanted to take a look at just the Borders in a time series idenifying the trend in the crude rate of cancer per 100,000.

#### - As we can see the trend clearly indicates an increase over time for All cancer types from 1995 to 2019

<br>

# Breakdowns by Cancer Type

<br>

#### - Next I wanted to identify the main drivers in this overall trend increase to get a better understanding of which specific types of cancer were driving the steady increase

#### - Plotting this out on a using the facet wrap gave too many graphs to be able to discern any trends. So I had to look about things in another way. Afterwards I figured out the solution was to make the plot into a function and then pass the function through plot_grid where you can select the plots shown.

<br>

```{r, warning=FALSE, message = FALSE}
table <- incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>%
  group_by(cancer_site) %>%
  summarise(percent_change_in_incidences = (incidences_all_ages/lag(incidences_all_ages, n = 24) - 1) *100,
            absolute_change_in_incidences = diff(incidences_all_ages, lag = 24) ) %>%
  arrange(desc(percent_change_in_incidences)) %>%
  head(21)

knitr::kable(table)
```

<br>

#### - I was able to calculate the absolute change and percentage change between 1995 and 2019, by using a summarise function. I then arranged the data by percent change and chose the top 20 values in terms of percent change to take a look at closer.

<br>

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Liver and intrahepatic bile ducts")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free") 
```

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Multiple myeloma and malignant plasma cell neoplasms")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Oropharyngeal cancers")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c( "Kidney")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Basal cell carcinoma of the skin")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Non-melanoma skin cancer")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Pancreas")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Lip, oral cavity and pharynx")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

```{r}
incidence_clean %>%
  filter(health_board_name == "Borders") %>%
  filter(sex == "All") %>% 
  filter(cancer_site == c("Squamous cell carcinoma of the skin")) %>%
  ggplot() +
  aes(x = year, y = incidences_all_ages, fill = cancer_site) +
  geom_col() +
  theme(legend.position = "none") +
  facet_wrap(~ cancer_site, scales = "free")
```

#### - From all 20 graphs I was able to pick out 9 cancer sites which showed consistent trend increases in incidences in the Borders. Moving forward, it would be preferable to do a linear regression to obtain the gradient, which would inform the expected rate increase that NHS Borders could expect year on year - for each group of cancer

#### - This could be very helpful to supply chain teams within hospitals who need to provision the treatments for these types of cancer.

#### - It would also inform hospital management and specialists on how they would be able to manage say, increased chemotherapy, for these expected patients.

#### - This could inform NHS Borders public health policy and messaging. There may be increased cancer awareness and early detection targeting these cancer types - informed by specialists and consultants - with what to look out for.

<br>

# Continuation:

<br>

#### - For the future, could use the age range statistics to inform what ages and sexs need targeting for messaging

#### - Breakdowns of Borders vs. Scotland vs. Europe