---
title: "Lists and Matrices"
output: html_notebook
---

```{r}

numbers <- 1:5

strings <- c("I", "am", "he")

booleans <- c(TRUE, FALSE)

class(numbers)

# These are all atomic vectors. vectors with the same type of value within

```
# Create them with `list()`

# back-ticks above render as a code format when knitted

```{r}

list("True", TRUE, 12, c(1,2,3), list("david", 2))

```

`as.list()`

```{r}
# Not in the tidyverse anymore! as denoted by the dot

as.list(numbers)
```

```{r}
x <- list(1, "a", TRUE, c(1,2,3))

class(x[[2]]) # Character

class(x[2]) # List

# Feeds back the position 2. 
```

`[]` will always return the same data structure

## Naming Lists

```{r}
named_list <- list(a = "I love RStudio", b = 1:10)

named_list
```

```{r}
named_list$b
```

```{r}

named_list[["a"]]

```

```{r}
returned_value <- "a"

named_list[[returned_value]]
```

```{r}

named_list[[2]]

```

## Nested Lists

You can put a list inside another list(inside another list!)

```{r}

student_info <- list(
                "Agnes" = list(height = 1.8, shoe_size = 8L),
                "Bobby" = list(height = 2.1, gender = "Male"),
                "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info

student_info$Bobby$height
```

```{r}

student_info$Agnes$shoe_size

# Here we are accessing items within lists within lists

```

```{r}

student_info[["Agnes"]][["shoe_size"]]

```

```{r}

student_info[[1]][[2]]

```

Task - 5 minutes

Make a named list that contains a numeric vector and a character vector.

Access the second element of your list using double square brackets [[]]. What is it's class?

Access the second element of your list using single square brackets []. What is it's class?

Access the second element of your list by name. What is it's class?

```{r}

# Question 1

example_list <- list(a = c(1,4,7,9,11), b =  c("J","M","L","R","P"))

example_list

# Question 2

example_list[[2]]

class(example_list[[2]])

# Question 3

example_list[2]

class(example_list[2])

# Question 4

example_list[["b"]]

class(example_list[["b"]])

# Additonal Q

class(example_list$b)

```

# Matrices

# A data structure that holds rectangular data

```{r}
matrix(1:6, nrow = 2, ncol = 3)
```

```{r}
matrix(1:6, nrow = 2, ncol = 3, byrow = TRUE)

# the "byrow = TRUE" argument fills by row first
```

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
```

```{r}
class(m)
```

# Matrices have to hold all values of the same type. 

```{r}
as.data.frame(m)
```
# Selecting element in a matrix

```{r}
m[1,3]

# Select value in the 1st row and 3rd column
```

```{r}
m[1,]

# Select the first row

m[,3]

# Select the third column
```

# You can select multiple rows or columns by passing a vector

```{r}

m[, c(1,3)]

# Returns everything from the first and third column

```

# Perform calculations on the entire matrix

```{r}
m * 5
```

```{r}
new_matrix <- matrix(c(1,2,5,745,45,5634), nrow=2, ncol=3)

new_matrix
```

```{r}
library(dplyr)
library(magrittr)

numers <- iris %>% select(-Species)

# if it's all the same data type

numers * 3

# access the first row of a df

numers[1, ]
```

## Working with rows and columns

- want to get the mean average of a column of a matrix

`colMeans()`

# NB: As there are no unscores in snakecase, we are not in the tidyverse

```{r}
colMeans(m)
```
# Other convenience functions:

`rowSums`
`colSums`
`rowMeans`
`colMeans`

```{r}

rowSums(m)

colSums(m)

rowMeans(m)

colMeans(m)

```

# These functions also work on data frames as long at the columns are all the same type.

```{r}
colMeans(numers)

# Spits a vector 
```

```{r}

```

##Task - 5 minutes
#Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 to 3.
#Find the element in the 2nd row, 3rd column.
#Check that all the columns are identical by checking that their sums are the same.

```{r}
the_matrix <- matrix(1:3, ncol = 4, nrow = 3)

the_matrix

the_matrix[2,3]

colSums(the_matrix)
```

