---
title: "Control Flow"
output: html_notebook
---

## If Statements

```{r}
if(condition) {
  #code that runs if the condition is met
}
```

# Conditional Operators

`==` is equal to
`!=` not equal to
`>` greater than
`<` less than
`%in%` is in the set

```{r}
if (5>3) {
  print("yes")
}
```

```{r}
x <- 7
y <- 8
if (x <= y) {
  print("it's less than or equal to y")
}
```

Write an if statement that prints 'Yes that divides by 2' if a number x is divisible by 2 (hint: use %% 2 == 0 as your expression), try it out for different values of x.

```{r}
x <- 11

if (x %% 2 == 0) {
  print("Yes that divides by 2")
}
```

## Nested Ifs

# We can write more complicated if statements

```{r}

if(10 != 5){
  if(2 == 0){
    print("the second condition is met")
  }
  print("the first condition is met")
}
```

```{r}

x <- 10
y <- 2

if (x > y) {
  print("Start of outer if block")

  if (x %% y == 0) {
    print("Inner if block 1")
  }

  if (y %in% c(1:5)) {
    print("Inner if block 2")
  }

  print("End of outer if block")
}

print("After if statement")
```

## If . . . Else

```{r}
if (condition) {
  # do something
} else {
  # do something else
}
```

```{r}
if(rain) {
  wear_wellies()
} else {
  wear_wellies()
}
```

```{r} 
if (4 < 5) {
  print("yes, its less")
} else {
  print("no, its not")
}
```

```{r}
if(2 %in% 50:100) {
  print("its in the set")
} else{
  print("its not in the set")
}
```

# Taskles

#Edit your code from the first exercise so that it prints ‘Yes that divides’ when x is divisible by 2, and prints ‘No that does not divide’ otherwise.

```{r}
x <- 9

if (x %% 2 == 0) {
  print("yes that divides")
} else{
  print("no that does not divide")
}
```

# Task

```{r}


if_else_function <- function(x, y){
  if (x %in% c(3, 4, 5)){
    
    print("x is in our list")
    
    if (x == 5){
      
      print("x is exactly 5")
      
    }
    
  } else {
    
    print("Start of else statement")
    
    if (y %% x == 0){
      
      print("x divides y")
    
    } else{
      
      print("x does not divide y")
    }
    
    print ("End of else statement")
    
  }
}

if_else_function(5, 10)
```

# Loops

## For Loops

# allow you to perform an action repeatedly

# R for loop syntax

```{r}
for(item in basket) {
  hang_up(item)
}
```

# Let's write a for loop to print the following statements.

```{r}
print("The day is Monday")
print("The day is Tuesday")
print("The day is Wednesday")
print("The day is Thursday")
print("The day is Friday")
```

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for (x in weekdays) {
  phrase <- paste("The day is", x, "oooweee", sep = " ")
  print(phrase)
}
```

Task that Loop!

Change the code below to use a for loop:

print(1)
print(2)
print(3)
print(4)
print(5)

```{r}
numbers <- 1:5

for (x in numbers) {
  phrase <- paste("print(", x, ")", sep = "")
  print(phrase)
}

# Actually what David wanted

numbers <-  c(1, 2, 3, 4, 5)

for (num in numbers) {
  print(num)
}
```

## While Loops

while (some condition) {
  do something
}

while (drive_filled_with_snow) {
 clear_drive()
}

### A more R example

```{r}

# set a beginning value
current_value <- 1
# set a stop value
stop_value <- 10

while(current_value < stop_value) {
  print(current_valu)
}

# will print 1 forever

# set a beginning value
current_value <- 1
# set a stop value
stop_value <- 10

while(current_value < stop_value) {
  print(current_value)
  # introduce a counter/ something to change the condition
  current_value <- current_value + 1
}
```

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for (i in 1:length(weekdays)) {
  print(i)
  phrase <- paste("The day is", weekdays[i], "oooweee", sep = " ")
  print(phrase)
}

```
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for (i in 1:length(weekdays)) {
  plural <- paste0(weekdays[i], "s")
  print(plural)
}

# We go through for the length of this weekdays vector, then just adding an s to the end of each bit within weekdays and printing

```

## Nested Loops

# Loops within loops

```{r}
# floor rounds to nearest whole number

#rnorm generates a small random number

my_matrix <- replicate(5, floor(rnorm(5) *100))

my_matrix
```
for row in rows{
 for col in cols{
  print(my_matrix[row, column])
 }
}
 
```{r}
n_rows <- nrow(my_matrix)
n_cols <- ncol(my_matrix)

for(row in 1:n_rows) {
  for(col in 1:n_cols){
    print(my_matrix[row, col])
  }
}

# This would work for a matrix and a data frame
```

# Taskles

```{r}
product_matrix <- matrix(nrow = 5, ncol = 5)
product_matrix

# This iterates through every row of the matrix
for (x in 1:nrow(product_matrix)){
# This iterates  through every column of the matrix
  for (y in 1:ncol(product_matrix)){
    print(x)
    print(y)
    product_matrix[x, y] <- x * y
  }
}

# Need to get the actual answer from the notes

```

```{r}
# Create a numeric vector
my_vector <- c(1, 4, 6, 2, 4)

# Create a temporary numeric vector (l = 5)
temp_vector <- numeric(length(my_vector))

for (i in 1:length(my_vector)) {
  temp_vector[i] <- my_vector[i] + 1
}
 # Overwrite initial vector
my_vector <- temp_vector

my_vector
```

```{r}
my_vector
my_vector + 1
```

## The Apply Family of Functions

`apply()`
`lapply()`
`sapply()`
`purr::map()`

```{r}
lapply(iris, class)

# sapply will try to return the simplest data structure it can (vectors)

#lapply will usually return a list
```

# Taskles

#Using the starwars list from the CodeClanData package, do the following:
#Print the name for every character in the list You can use for loops or apply

```{r}
for (name in starwars$name) {
  print(name)
}

# OR

sapply(CodeClanData::starwars, '[[', "name")

# OR

starwars[[1]]
```

