---
title: "two_sample_tests"
author: "Jonny Nelson"
date: "09/12/2021"
output: html_document
---

```{r}
# Be wary of loading janitor after tidyverse

library(infer)
library(janitor)
library(prob)
library(tidyverse)
```

# Apartments in Nice and Algarve

# Reading in the data

```{r}
algarve_data <- read_csv("3_two_sample_tests/data/algarve.csv") %>%
  clean_names()

nice_data <- read_csv("3_two_sample_tests/data/nice.csv") %>%
  clean_names()
```

# Apend the data-sets

```{r}
apart_prices <- bind_rows(nice_data, algarve_data, .id = "location") %>%
  mutate(location = if_else(location == "1",
                            "nice",
                            "algarve")) %>%
  select(-x1)
```

# Visualise the new data

```{r}
apart_prices %>%
  group_by(location) %>%
  summarise(n = n())
```

```{r}
apart_prices %>%
  ggplot(aes(x = location, y = price)) +
  geom_boxplot()
```

# Null Distribution

```{r}
null_distributions <- apart_prices %>%
  specify(price ~ location) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 10000, type = "permute") %>%
  calculate(stat = "diff in means",
            order = c("algarve", "nice"))

null_distributions
```

# Observed Stat

```{r}
observed_stat <- apart_prices %>%
  specify(price ~ location) %>%
  calculate(stat = "diff in means",
            order = c("algarve", "nice"))
```

# Visualising the Null Distribution

```{r}
null_distributions %>%
  visualise(bins = 30) +
    shade_p_value(obs_stat = observed_stat,
                  direction = "right")
```

# Calculating the p_values

```{r}
p_value <- null_distributions %>%
  get_p_value(obs_stat = observed_stat,
             direction = "right")

p_value
```


# Florence and Corfu


# Reading in the data

```{r}
florence_data <- read_csv("3_two_sample_tests/data/florence.csv") %>%
  clean_names()

corfu_data <- read_csv("3_two_sample_tests/data/corfu.csv") %>%
  clean_names()

corfu_data %>%
  summarise(mean = mean(price))

florence_data %>%
  summarise(mean = mean(price))
```

# Apend the data-sets

```{r}
apart_prices <- bind_rows(florence_data, corfu_data, .id = "location") %>%
  mutate(location = if_else(location == "1",
                            "florence",
                            "corfu")) %>%
  select(-x1)
```

# Visualise the new data

```{r}
apart_prices %>%
  group_by(location) %>%
  summarise(n = n())
```

```{r}
apart_prices %>%
  ggplot(aes(x = location, y = price)) +
  geom_boxplot()
```

# Null Distribution

```{r}
null_distributions <- apart_prices %>%
  specify(price ~ location) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 70000, type = "permute") %>%
  calculate(stat = "diff in means",
            order = c("florence", "corfu"))

null_distributions
```
# Observed Stat

```{r}
observed_stat <- apart_prices %>%
  specify(price ~ location) %>%
  calculate(stat = "diff in means",
            order = c("florence", "corfu"))
```

# Visualising the Null Distribution

```{r}
null_distributions %>%
  visualise(bins = 30) +
    shade_p_value(obs_stat = observed_stat,
                  direction = "left")
```

# Calculating the p_values

```{r}
p_value <- null_distributions %>%
  get_p_value(obs_stat = observed_stat,
             direction = "left")

p_value

# Null Hypothesis: population mean of florence - population mean of corfu is 0

# Altnerate Hypothesis: florence sign lower than corfu

# population mean of florence - population mean of corfu < 0 
```
# 2-Sample Test: Paired/Dependent

```{r}
books <- read_csv("3_two_sample_tests/data/ucla_textbooks_f18.csv")
```

```{r}
glimpse(books)
```

```{r}
books_diff <- books %>%
  mutate(diff_new = bookstore_new - amazon_new) %>%
   dplyr::filter(!is.na(diff_new))

books_diff %>%
  ggplot(aes(x = diff_new)) +
  geom_histogram(col = "white")
  
```
# Null Hypothesis

H0: mu(diff_new) = 0
HA: mu(diff_new) != 0

# Null Distribution

```{r}
null_distribution <- books_diff %>%
  specify(response = diff_new) %>%
  hypothesise(null = "point", mu = 0) %>%
  generate(reps = 50000, type = "bootstrap") %>%
  calculate(stat = "mean")
```

# Null Distribution

```{r}
null_distribution %>%
  visualise(bins = 100) +
  shade_p_value(obs_stat = observed_stat,
                direction = "both")
```


```{r}
observed_stat <- books_diff %>%
  specify(response = diff_new) %>%
  calculate(stat = "mean")

observed_stat
```

# Calcaulte p_value

```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat,
              direction = "both")

p_value

# We are able to reject the null hypothesis as the p_value is below 5%, therefore we fail to reject the alternative hypothesis that the diff_mean new book price -> significantly different.
```


# Testing the Difference in Proportions

```{r}
poll <- read_csv("3_two_sample_tests/data/election_data.csv")

head(poll)
```

# Visualise the data we have

```{r}
ggplot(poll, aes(x = vote, fill = gender)) +
  geom_bar(position = "fill")+
  coord_flip()
```

# Difference in Proportion in the observed statistic

```{r}
observed_stat <- poll %>%
  specify(gender ~ vote,
          success = "woman") %>%
  calculate(stat = "diff in props",
            order = c("obama", "romney"))

observed_stat
```

```{r}
null_distributions <- poll %>%
  
  specify(gender ~ vote,
          success = "woman") %>%
  
  hypothesise(null = "independence") %>%
  
  generate(reps = 1000,
           type = "permute") %>%
  
  calculate(stat = "diff in props",
            order = c("obama", "romney"))

null_distributions
```

# Visualise the Null Distribution

```{r}
null_distributions %>%
  visualise(bins = 30) +
  shade_p_value(obs_stat = observed_stat,
                direction = "both")
```

```{r}
p_value <- null_distributions %>%
  get_p_value(obs_stat = observed_stat,
              direction = "both")

p_value
```

