---
title: "distributions_and_measures_of_spread"
author: "Jonny Nelson"
date: "07/12/2021"
output: html_document
---

```{r}
library(tidyverse)
library(janitor)
```

# Measures of Spread

## Range - Minimum to the Maximum

```{r}
jobs <- read_csv("2_distributions_measures_of_spread/data/TyrellCorpJobs.csv") %>%
  clean_names()

jobs %>%
  summarise(max = max(salary),
            min = min(salary),
            
            min_r = range(salary)[1],
            max_r = range(salary)[2],
            range = max(salary) - min(salary))

range(jobs$salary)

# range is a vector of 2, can then select these above.

diff(range(jobs$salary))

# gives the range
```

```{r}
# positions together

jobs %>%
  ggplot() +
  geom_histogram(aes(x = salary), col = "white", bins = 25)
```

```{r}
jobs %>%
  ggplot() +
  geom_histogram(aes(x = salary), col = "white", bins = 25) +
  facet_wrap(~position)

# Separating by position
```

```{r}
jobs %>%
  group_by(position) %>%
  summarise(range = diff(range(salary)))
  
```

## A potential problem with using range as a measure of spread: very strongly affected by outliers.

# Quartiles and IQR

# Q1 - 25% fence post split
# Q2 - 50% post
# Q3 - 75% post

--|--|--|--

```{r}
# Quartiles for the jobs data

jobs %>%
  group_by(position) %>%
  summarise(Q1 = quantile(salary, 0.25),
            Q2 = quantile(salary, 0.50),
            Q3 = quantile(salary, 0.55),
            )
```

```{r}
# quartiles on the vis
iqr <- jobs %>% 
  group_by(position) %>% 
  summarise(
    Q1 = quantile(salary, 0.25),
    Q2 = quantile(salary, 0.5),
    Q3 = quantile(salary, 0.75)
  ) %>% 
  pivot_longer(Q1:Q3, names_to = "fencepost", values_to = "value")

jobs %>% 
  ggplot() +
  geom_histogram(aes(x = salary), col = "white", bins = 25) +
  geom_vline(xintercept = iqr$value, linetype = "dashed", colour = "blue") +
  facet_wrap(~position)
```

# Interquartile Range Mayte

```{r}
jobs %>%
  group_by(position) %>% 
  summarise(
    Q1 = quantile(salary, 0.25),
    Q2 = quantile(salary, 0.5),
    Q3 = quantile(salary, 0.75),
    IQR = Q3 - Q1,
    # R also has an iqr function
    IQR_fun = IQR(salary)
  )
```

# The five figure summary

# Min, Q1, Q2 (Median), Q3, Max

```{r}
library(skimr)

# skim is a more sophisticated summary
```
```{r}
jobs %>%
  select(-x1) %>%
  group_by(position) %>%
  skim()
```

# Box Plotsz


```{r}
jobs %>%
  ggplot() +
  aes(x = salary, y = position) +
  stat_boxplot(geom = "errorbar", width = 0.5) +
  geom_boxplot()

# whiskers are typically positioned 1.5 IQR above and below Q3 and Q1 resp'
```

# Skew in Box Plots

```{r}
heavily_right_skewed <- read_csv("2_distributions_measures_of_spread/data/heavily_right_skewed.csv")

heavily_right_skewed %>%
  ggplot() +
  aes(x = x) +
  stat_boxplot(geom = "errorbar", width = 0.5) +
  geom_boxplot()
```

```{r}
heavily_right_skewed %>%
  ggplot() +
  geom_histogram(aes(x=x), colour = "white", bins = 30)
```

## Taskles 

```{r}
# Find skewness of data above

library(e1071)

jobs %>%
  group_by(position) %>%
  summarise(skewness = skewness(salary, type = 1))
```

# Variance $s^2$

## Variance - measure of how far values deviate from the mean.

## How to calculate the variance 

### 1. get the difference between each value and the mean
### 2. Sum
### 3. divide by the total number of values

```{r}
jobs %>%
  group_by(position) %>%
  summarise(variance = var(salary),
            standard_deviation = sd(salary),
            standard_deviation_01 = sqrt(var(salary)))

# Blammo!
```

## Standard Deviation - a more interpertable measure of spread
