---
title: "Untitled"
author: "Jonny Nelson"
date: "07/12/2021"
output: html_document
---
```{r}
# Order is important

library(prob)
library(tidyverse)
```

# Distributions: A rigorous definition

## A distribution is a list of all of the possible outcomes of a random variable along with either their corresponding frequency (frequency dist') or probability values (probability dist')

## Less rigorous: distributions tell us how common / how likely various outcomes are.

```{r}
s_three_coins <- tosscoin(3)

s_three_coins
```

# Taskles

```{r}

s_three_coin_p <- s_three_coins %>%
  mutate(label = str_c(toss1, toss2, toss3)) %>%
  mutate(prob = 1/n())

s_three_coin_p %>%
  ggplot() +
  geom_col(aes(x= label, y = prob, colour = "steel blue"))

# uniform distribution: all the bars are the same height, indicating that all probabilites are exactly the same

```
  
```{r}
s_three_coin_p %>%
  mutate(n_heads = str_count(label, "H")) %>%
  group_by(n_heads) %>%
  summarise(prob = sum(prob)) %>%
  ggplot() +
  geom_col(aes(x= n_heads, y = prob, colour = "steel blue"))
```
## Distibution is no longer uniform, but is still discrete

## General properties of discrete distributions

## y - axis = probability or frequency 
## x - axis = outcomes

## the integral of the y-axis must always == 1



# Measures of Centrality 

```{r}

library(janitor)
library(lubridate)

air_con_sales <- read_csv("1_distributions_measures_of_centre/data/AirConSales.csv") %>%
  clean_names()

# Wrangle date column
air_con_sales <- air_con_sales %>%
  mutate(date = mdy(date))

```

```{r}

# We have measurements for each day, but we want to look at how likely it will be to sell a certain number of units (air conditioners)

# Generate a frequency table

sales_freq_table <- air_con_sales %>%
  tabyl(units_sold) 

sales_freq_table

```


```{r}
sales_freq_table %>%
  ggplot() +
  geom_col(aes(x = units_sold, y = n))
```

## mean == sum of values/number of all values
## median == 50% of data beneath, 50% of data above
## mode == most frequent value

## R does not have a mode function

```{r}
# Function to get the mode

get_mode <- function(data){

  tabled_data <- table(data)
  table_names <- names(tabled_data)
  
  return(as.numeric(table_names[tabled_data == max(tabled_data) ]))
  
}
```

## Lets calculate these values for out distibution

```{r}
air_con_sales %>%
  summarise(mean_daily_sales = mean(units_sold),
            median_daily_sales = median(units_sold),
            mode_daily_sales = get_mode(units_sold))
```

## Extremes inform the means

# Measures of centre and outliers

## If you suspect or observe outliers in your data, be careful which measure of centre you use.

## Bimodal - two peaks
## Unimodal - one peak


## Skewness - assymetry of a distribution


## left-skew - left tailed

## the mean < median < mode


## right-skew - left tailed

## the mean > median > mode

```{r}
library(e1071)
```

```{r}
left_skewed <- read_csv("1_distributions_measures_of_centre/data/leftskew.csv")
right_skewed <- read_csv("1_distributions_measures_of_centre/data/rightskew.csv")

left_skewed %>%
  summarise(skewness = skewness(x, type = 1))

# left-skew ~ -1

right_skewed %>%
  summarise(skewness = skewness(x, type = 1))

# right_skew ~ 0.7

# Both have a direction and magnitude
```

# Taskles 

```{r}
air_con_sales %>%
  summarise(skewness = skewness(units_sold, type = 1))

# Air Con Sales are heavily right skewed
```



