---
title: "Lesson 2"
author: "Jonny Nelson"
date: "08/12/2021"
output: html_document
---

# Confidence Intervals

## Learning Objectives

### What are confidence intervals
### understand and perform bootstrap resampling
### use the infer package to do this
### use infer to calculate confidence intervals

```{r}
library(e1071)
library(janitor)
library(infer)
library(tidyverse)

telecoms_data <- read_csv("1_sampling_distributions_standard_errors/data/telecomms_churn.csv") %>%
  clean_names
```

```{r}
sample_200 <- telecoms_data %>%
  rep_sample_n(
    size = 200, 
    reps = 1
  )
```

```{r}
# Statistics for the sample of 200

summary_sample_200 <- sample_200 %>%
  
  # Currrently grouping by the replicate, need to ungroup()
  
  ungroup() %>%
  summarise(
    mean_monthly_charges = mean(monthly_charges), 
    mean_tenure = mean(tenure),
    prop_churn = mean(churn == "Yes")
  ) 
summary_sample_200
```

```{r}

# statistics for the population

 telecoms_data %>% 
  ungroup() %>%
  summarise(
    mean_monthly_charges = mean(monthly_charges), 
    mean_tenure = mean(tenure),
    prop_churn = mean(churn == "Yes")
  ) 
```

# Confidence Intervals - a range where we can say that with a degree of confidence the true value exsists

### e.g 95% confidence level

### 31.42 is the mean(tenure) whereas the real value is 32.37

## The Confidence Interval Formula

## = sample mean + / - sample sd / sqrt(sample size)


## For 95% confidence level, if we took 100 samples, calculating point estimates (mean(tenure)) and confidence intervals for each, we would expect 95 of the determined confidence intervals to contain the true value of the population parameter. 

# CI is the range of values around a point estimate where we are confident the population value exsists

# CI are proportional to the standard error

## typically larger sample size - narrow confidence intervals



# Boot-Strapping

## In practice we only have one or a few samples to work with



```{r}
de11 <- c("jonny", "louise", "malcolm", "ricardo", "pui")

sample(de11, replace = TRUE)
```

```{r}
sample_200 <- read_csv("4_confidence_intervals/data/telecomms_churn_200_sample.csv") %>%
  clean_names()

head(sample_200)
```

```{r}
# Point estimate is mean(tenure)

bootstrap_resample <- sample_200 %>%
  # resample 5000 times
  rep_sample_n(size = 200,
               reps = 5000,
               replace = TRUE) %>%
  summarise(
    mean_tenure = mean(tenure)
  )

bootstrap_resample
```

## Taskles

```{r}
bootstrap_resample %>%
  ggplot() +
  aes(x = mean_tenure) +
  geom_histogram(col = "white", fill = "steel blue") +
  labs(x = "mean_tenure")
```


```{r}
standard_error <- bootstrap_resample %>%
  dplyr::summarise(se_mean_tenure = sd(mean_tenure))

standard_error

# Confidence Interval Formula

mean(bootstrap_resample$mean_tenure) - (sd(bootstrap_resample$mean_tenure)) / sqrt(200)
```

# se ~ 1/sqrt(sample_size)
# ci ~ 1/sqrt(sample_size)

# bootstrapping is just computer-based accuracy metric, not creating ficticous data

# Confidence Intervals from Bootstrapping Distribution

```{r}
CI_95 <- bootstrap_resample %>%
  
  summarise(
    mean = mean(mean_tenure),
    
    # 95% Confidence Interval
    
    lower_bound = quantile(mean_tenure, probs = 0.025),
    upper_bound = quantile(mean_tenure, probs = 0.975)
    
  )

CI_95

# mean tenure = 32.62
# 95% confident the true value of the mean is within 29.20 - 36.07
```

```{r}
# 99% Confidence Interval

CI_99 <- bootstrap_resample %>%
  
  summarise(
    mean = mean(mean_tenure),
    
    # 99% Confidence Interval
    
    lower_bound = quantile(mean_tenure, probs = 0.01),
    upper_bound = quantile(mean_tenure, probs = 0.99)
    
  )

CI_99

# mean tenure = 32.62
# 95% confident the true value of the mean is within 28.62 - 36.65
```

# Workflow - Specify, Generate, Calculate

```{r}
# 1. Derive bootstrap distribution
# 2. Get confidence interval
# 3. Visualise our CI
# 4. Get our point estimate
```

```{r}
# 1. Derive bootstrap distribution

infer_resample <- sample_200 %>%
  specify(response = tenure) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

infer_resample
```

```{r}
# 2. Derive confidence level

infer_ci_95 <- infer_resample %>%
  get_ci(level = 0.95, type = "percentile")

infer_ci_95
```


```{r}
# 3. Visualise Distribution and CI

infer_resample %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_95)
```

```{r}
# 4. Get our point estimate

mean_infer <- infer_resample %>%
  summarise(mean = mean(stat))

mean_infer
```

```{r}
sample_300 <- read_csv("4_confidence_intervals/data/telecomms_churn_300_sample.csv") %>%
  clean_names()
```

```{r}
# 1. Derive bootstrap distribution

infer_resample_300 <- sample_300 %>%
  specify(response = tenure) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

infer_resample_300
```

```{r}
# 2. Derive confidence level

infer_ci_95_300 <- infer_resample_300 %>%
  get_ci(level = 0.95, type = "percentile")

infer_ci_95_300
```


```{r}
# 3. Visualise Distribution and CI

infer_resample_300 %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_95_300)
```

```{r}
# 4. Get our point estimate

mean_infer_300 <- infer_resample %>%
  summarise(mean = mean(stat))

mean_infer_300
```

# CI for a proportion

```{r}
sample_200 %>%
  distinct(churn)
```

```{r}
bootstrap_resample_churn <- sample_200 %>%
  specify(response = churn, success = "Yes") %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "prop") 

bootstrap_resample_churn
```

```{r}
# Get the CI boundaries

churn_ci_95 <- bootstrap_resample_churn %>%
  get_ci(level = 0.95, type = "percentile")

churn_ci_95
```

```{r}
# Visualise 

bootstrap_resample_churn %>%
  visualise(bins = 30) +
  shade_ci(endpoints = churn_ci_95)
```

```{r}
mean_200_churn <- bootstrap_resample_churn %>%
  summarise(point_estimate = mean(stat))

mean_200_churn

# Number of customers that do churn is 0.254 with a 95% confidence, that the true proportion is between 0.195 - 0.315
```

