---
title: "spatial_analysis"
author: "Jonny Nelson"
date: "14/12/2021"
output: html_document
---

# Spatial Data

```{r}
library(sf)
library(tidyverse)
library(janitor)
```

# Spatial Data is special - while encoded as numbers - these numbers actually have a special meaning

spatial vectors (non R-specific)

## 3 main types

- point data -> POINT (1,4)
- line -> LINE STRING ((1,4)(4,5))
- polygon -> POLYGON ((x,y)(x,y)(x,y)(x,y)(x,y))

## Examples for 3 main types

- points (stations, wells)
- lines (roads, rivers)
- polygon (council areas, lakes)

## lat long columns - or load shapefiles

```{r}
library(sf) # sf = simple features
```

# a feature is just a geometry

```{r}
north_carolina <- st_read(system.file("shape/nc.shp", package = "sf"))
```

```{r}
north_carolina %>%
  as_tibble()
```

```{r}
nc_geo <- st_geometry(north_carolina)

nc_geo[[1]]
```
```{r}
class(north_carolina)
```

# Plotting geometries

```{r}
plot(north_carolina["AREA"])
```

```{r}
# `st_as_sf()` - convert non-sf objects to sf data frames
```

```{r}
# first row, "AREA" column

plot(north_carolina[1, "AREA"])
```

```{r}
plot(north_carolina["CNTY_ID"])
```

# ggplot and sf

`geom_sf()`

```{r}
north_carolina %>%
  ggplot() +
  geom_sf(aes(fill = SID74), colour = "black") +
  theme_bw()
```

# Taskles

```{r}
north_carolina %>%
  ggplot() +
  geom_sf(aes(fill = CNTY_ID), colour = "red", size = 0.1) +
  scale_fill_continuous(type = "viridis") +
  theme_void()
```

```{r}
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
```

# Using rnaturalearth package we can inmport boundaires of countries at various levels

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")

class(world)
```

```{r}
world %>%
  as_tibble() %>%
  head(5)
```

```{r}
world %>%
  ggplot() +
  geom_sf() +
  labs(x = "longitude", y = "latitude", title = "World Map")
```

```{r}
world %>%
  ggplot() +
  geom_sf(aes(fill = pop_est)) +
  scale_fill_viridis_c(trans = "sqrt")
```

```{r}
world %>%
  ggplot() +
  geom_sf(aes(fill = gdp_md_est)) +
  scale_fill_gradient2()
```

```{r}
world %>% 
  filter(continent != "Antarctica") %>%
  ggplot() +
  geom_sf(aes(fill = continent))
```

with the world dataframe, we can filter out different countries

```{r}
# unique(world$sovereignt)

country_italy <- world %>%
  filter(name == "Italy")

# Plot just Italy

country_italy %>%
  ggplot() +
  geom_sf() +
  labs(x = "Longitude", y = "Longitude")
```

# Pick a country filter and plot it

```{r}
unique(world$sovereignt)

country_liechtenstein <- world %>%
  filter(name == "Liechtenstein")

# Plot just Liechtenstein

country_liechtenstein %>%
  ggplot() +
  geom_sf() +
  labs(x = "Longitude", y = "Longitude")
```

## Zooming in on a particular parts of the world

- We can subset our graph by limiting the x and y range  `coord_sf()`

```{r}
world %>%
  ggplot() +
  geom_sf() +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```

# Add informative labels

- we nedd to tell ggplot where top put the labels 
- we need to calucalte where we want to put the labels
(centre of the feature)
- we need to calculate the centres (centroids)

```{r}
world %>%
  mutate(centres = st_centroid(st_make_valid(geometry))) %>%
  as_tibble() %>%
  select(name, centres)
```
```{r}
# mutated_geometries

world_with_centres <- world %>%
  mutate(centres = st_centroid(st_make_valid(geometry))) %>%
  mutate(lat = st_coordinates(centres)[,1],
         long = st_coordinates(centres)[,2])
```

```{r}
# centre_geometries

centre_geo <- world %>%
  mutate(centres = st_centroid(st_make_valid(geometry))) %>%
  select(centres)
```

```{r}
world_with_centres %>%
  ggplot() +
  geom_sf() +
  geom_text(aes(x = lat, y = long, label = name), colour = "darkblue", fontface = "bold", check_overlap = TRUE, size = 2.5) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) 
```

# We can add additonal information using annotate

- add one that says: Gulf of Mexico

```{r}
world_with_centres %>%
  ggplot() +
  geom_sf() +
  geom_text(aes(x = lat, y = long, label = name),
            colour = "darkblue",
            fontface = "bold",
            check_overlap = TRUE,
            size = 2.5) +
  annotate(geom = "text", x = -90, y = 26, label = "Gulf Of Mexico", size = 3, fontface = "bold") +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97),
           expand = FALSE) 
```

```{r}
world_with_centres %>%
  ggplot() +
  geom_sf() +
  geom_text(aes(x = lat, y = long, label = name, colour = "seagreen"),
            colour = "darkblue",
            fontface = "bold",
            check_overlap = TRUE,
            size = 2.5) +
  annotate(geom = "text", x = -90, y = 26, label = "Gulf Of Mexico", size = 3, fontface = "bold") +
  coord_sf(xlim = c(6,12), ylim = c(44,50),
           expand = FALSE) 
```

wgs84 (1984) - how do we put a sphere on a rectangle - need another argument to make it work for smaller long and lats.


```{r}
world_with_centres %>% 
  ggplot() +
  geom_sf(aes(fill = name), show.legend = F) +
  geom_text(aes(lat, long, label = name), size = 2.5) +
  annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico",
           size = 4, fontface = "italic",
           check_overlap = TRUE) +
  coord_sf(xlim = c(-110, -60), ylim = c(8, 35), expand = T) +
  theme_minimal() 
```

# interactive Maps with Leaflet

```{r}
library(leaflet)

leaflet() %>%
  addTiles() %>% # basemap
  addMarkers(lng = 174.768, lat = -36.852, popup = "The Birthplace of R")
```

# Can get some spatial data from the web-a-roo
- turn it into the correct format
- visualise it using leaflet

```{r}
library(jsonlite)
```

```{r}
colorado_data_url <- "https://data.colorado.gov/resource/j5pc-4t32.json?&county=BOULDER"
```

```{r}
head(readLines(colorado_data_url))
```

jsonlite package has alot of functions to help work with json data

json is like a list of lists in R - sometimes we need to recursively sift through to extract the relevant data

```{r}
colorado_water <- fromJSON(colorado_data_url) %>%
  jsonlite::flatten(recursive = TRUE)
```

```{r}
colorado_water %>%
  select(starts_with("location"))

# need to do some wrangling before the plotting
```

```{r}
colorado_water_clean <- colorado_water %>%
  select(-location.needs_recoding) %>%
  mutate(across(starts_with("location"), as.numeric)) %>%
  filter(!is.na(location.latitude), !is.na(location.longitude))
```

# Visualise it with leaflet

```{r}
colorado_water_clean %>%
leaflet() %>%
addTiles() %>%
addCircleMarkers(lng = ~location.longitude,
                 lat = ~location.latitude,
                 radius = ~amount/10)

# incidences of surface water in and around boulder colorado
```

# Clustering

Lets have a look at AddMarkers

```{r}
colorado_water_clean %>%
leaflet() %>%
addTiles() %>%
addMarkers(lng = ~location.longitude,
                 lat = ~location.latitude,
                 )
```

We can add clustering options

```{r}
colorado_water_clean %>%
leaflet() %>%
addTiles() %>%
addMarkers(lng = ~location.longitude,
           lat = ~location.latitude,
           clusterOptions = markerClusterOptions()
           )
```

# Leaflet in Shiny

```{r}
whiskey <-CodeClanData::whisky %>%
  rename(long = Latitude,
         lat = Longitude)

head(whiskey)
```

```{r}
whiskey %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lat = ~lat, lng = ~long, popup = ~Distillery)
```

