---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
tidyverse_logo()
```
```{r}
tidyverse::tidyverse_logo()
```

```{r}
devtools::install_github("codeclan/CodeClanData")
```

```{r}
devtools::install_github("codeclan/CodeClanData")
```

```{r}
class(CodeClanData::starwars)
```

```{r}
install.packages("devtools")
```

```{r}
devtools::install_github("codeclan/CodeClanData")
```

```{r}
library(CodeClanData)
```

```{r}
students
```
# Working with dplyr

```{r}
all_deaths
```
#Chekcing out Data

```{r}
# number of rows
nrow(all_deaths)
```
```{r}
# number of columns
ncol(all_deaths)
```
```{r}
# overall dimensions of our data
dim(all_deaths)
```

```{r}
# variable names
names(all_deaths)
```

```{r}
# check the first 10 rows

head(all_deaths, 15)
```

```{r}
# Check the tail end of the data

tail(all_deaths, 15)
```

```{r}
# Give an overview of the data

glimpse(all_deaths)
```

```{r}
#view the data in a table 

view(all_deaths)
```

```{r}
# view a short representation of the data

str(all_deaths)
```

```{r}
# print data into the console

all_deaths
```
#select

```{r}
# Using the select() function to obtain certain columns. Put the name of tib 1st

select(all_deaths, name, allegiances, gender, nobility, year_of_death)
```

```{r}
# exclude one column, just at a "-" symbol and the name of the column

select(all_deaths, -name)
```

```{r}
deaths_no_names <- select(all_deaths, -name) 
```

```{r}
# First Attempt

# Select all columns that relate to a book 

GOT_death_books <- select(all_deaths,-name, -allegiances, -year_of_death,
                          -book_of_death, -death_chapter,-book_intro_chapter, 
                          -gender, -nobility, )

# Select all the columns related to some kind of death 

GOT_related_to_some_kind_of_death <- select(all_deaths, book_of_death, death_chapter, year_of_death)
```

```{r}
# Alternative Solutions

# Select all columns that relate to a book 

GOT_death_books_solution <- select(all_deaths, 9:13)

GOT_death_books_solution_2 <- select(all_deaths, book1_GoT:book5_DwD)

GOT_death_books_solution_3 <- select(all_deaths, contains("book"))

# Select all the columns related to some kind of death 

Second_Question_Sol_1 <- select(all_deaths, contains("death"))

```

#filter()

```{r}

# Filter. "==" go through allegiances and pick only "Lannister" 

filter(all_deaths, allegiances == "Lannister")


# the "|" symbol allows to choose the OR function

filter(all_deaths, allegiances == "Lannister" | allegiances == "House Lannister")


#Filter all observations not in Lannister

filter(all_deaths, allegiances != "Lannister")


#Filter all deaths where the allegiances are in Lannister and House Lannister

filter(all_deaths, allegiances %in% c("House Lannister", "Lannister"))


```

```{r}
#Filter with numerics 

filter(all_deaths, year_of_death > 250)

```

```{r}
#Filter those who died in the year 250 or equal to

filter(all_deaths, year_of_death >= 250)
```

```{r}
filter(all_deaths, allegiances == "None")
```

```{r}
filter(all_deaths, book_intro_chapter >= 5 & book_intro_chapter)
```

#Task

#Find where the year_of_death is less than or equal to 299.

```{r}
filter(all_deaths, year_of_death <= 299)
```

#Find the females (gender is 0) who are not Lannisters

```{r}
filter(all_deaths, gender == 0 & allegiances != c("House Lannister", "Lannister"))
```

#Find just the data for the characters “Jon Snow”, “Daenerys Targaryen” and “Samwell Tarly”.

```{r}
filter(all_deaths, name == "Jon Snow" | "Daenerys Targaryen"| "Samwell Tarly")
```

```{r}
filter(all_deaths, name %in% c("Jon Snow", "Daenerys Targaryen", "Samwell Tarly"))
```

#arrange

```{r}
arrange(all_deaths, gender)
```

```{r}
arrange(all_deaths, desc(gender))
```

```{r}
#Sorting by book of death then death chapter, chronologically 

arrange(all_deaths, book_of_death, death_chapter)
```

#Task

#Arrange all_deaths by allegiances. What happens when you arrange by a character column?

```{r}
arrange(all_deaths, allegiances)

#Sorts by alphabetical order.
```

#Arrange all_deaths by allegiances and book_intro_chapter

```{r}
arrange(all_deaths, allegiances, book_intro_chapter)
```

#Arrange all_deaths by descending year_of_death

```{r}
arrange(all_deaths, desc(year_of_death))
```

#Mutate

```{r}
mutate(all_deaths, years_survived = year_of_death - 298) 
```

```{r}
mutate(all_deaths, book_of_death = book_of_death * 5)
```

```{r}
mutate(all_deaths, year_of_death_logical = is.na(year_of_death))

# If year_of_death is "N/A" then it is changed to TRUE
```


```{r}
mutate(all_deaths, book_of_death = as.character(book_of_death))

# Converts column "book_of_death" from numeric to character
```

```{r}
mutate(all_deaths, name = as.numeric(name))

# the names can't really be numeric. So an alert pops up saying it has been coerced
```

```{r}
mutate(all_deaths, year_of_death = sum(year_of_death, na.rm = TRUE))

#na.rm means N/A remove
```

```{r}
mutate(all_deaths, year_of_death = mean(year_of_death, na.rm = TRUE))
```


#Summarise Function

```{r}
summarise(all_deaths, n_males = sum(gender))
```

```{r}
deaths_grouped <- group_by(all_deaths, allegiances)

summarise(deaths_grouped, character_count = n())

#Created a new variable where the allegiances are grouped

#Summarized the new variable and done a character count on the grouped allegiances
```

```{r}
deaths_grouped <- group_by(all_deaths, nobility, gender)

summarise(deaths_grouped, character_count = n())
```


# Magrittr Pipe %>%

```{r}
deaths_by_allegiance <- all_deaths %>% 
  filter(!is.na(book_of_death)) %>%
  group_by(allegiances) %>%
  summarise(Number_of_Deaths_per_House = n()) %>%
  arrange(desc(Number_of_Deaths_per_House))

deaths_by_allegiance

```

# Pull Function

```{r}
avg_year_of_death <- all_deaths %>%
  summarise(av_value = mean(year_of_death, na.rm = TRUE))

avg_year_of_death
```

```{r}
all_deaths %>%
  mutate(death_later_than_av = year_of_death > avg_year_of_death)
```

```{r}
all_deaths %>%
  mutate(death_later_than_av = year_of_death > 299.1574)

#using a number in there, doesn't allow the coded to be flexiible when the data changes
```

```{r}
class(avg_year_of_death)
class(299.1574)
```

```{r}
av_year_of_death <- all_deaths %>%
  summarise(av_value = mean(year_of_death, na.rm = TRUE)) %>%
  pull()

av_year_of_death
```

```{r}
all_deaths %>%
  mutate(death_later_than_av = year_of_death > av_year_of_death)

#preferable to having a number hard coded in there. Makes the results more variable for when the data changes
```

