---
title: "R Notebook"
output: html_notebook
---

# regex = regular expression. This allowed code to be more flexible and allows us to view patterns.

```{r}
library(tidyverse)
```

```{r}
single_string <- "sting a"
strings <- c("string a", "string b", "string c", "string d", "striiiiiiing e")
```

```{r}
pattern <- "a"
```

```{r}
single_string %>%
  str_detect(pattern)
```
```{r}
strings %>%
  str_detect(pattern)
```

```{r}
pattern <- "str"
strings %>%
  str_detect(pattern)

# The pattern is "str", and then detecting across strings for this pattern
```

```{r}
pattern <- "[abc]"

# Find if there is an "a" OR a "b" OR a "c".

# "[]" this is Regex.

strings %>%
  str_detect(pattern)
```

```{r}
pattern <- "[a-z]" # look for letters between a - z

strings %>%
  str_detect(pattern)
```

# Regex is case sensitive

```{r}
pattern <- "[A-Z]" # look for letters between A - Z

strings %>%
  str_detect(pattern)

# No capitals so no returns TRUE
```

```{r}
pattern <- "[a-zA-Z]" # look for letters between a - c, A-Z

strings %>%
  str_detect(pattern)
```

```{r}
strings
```

```{r}
pattern <- "i{3}" # look for 3 consecutive lowercase "i's"

strings %>%
  str_detect(pattern)
```

# Task 1: 

- Write and test regular expressions which will match:
- Any string mentioning CodeClan
- Any string containing two consecutive capital letters
- Any string not including the word “castle” or “Castle”

```{r}
test_strings <- c(
  "We're having great fun learning about data analysis at CodeClan!",
  "CodeClan is at 37 Castle Terrace in Edinburgh, ClockWise in Glasgow, and Wasps Academy in Inverness",
  "Our cohort code starts with DE, DG or D",
  "You get a great view of the castle or the national conservatoire from just outside CodeClan's door"
)
```

```{r}
pattern <- "CodeClan"

test_strings %>%
  str_detect(pattern)

# "[Cc]ode[Cclan]" if you wanted either title or lowercase
```

```{r}
pattern <- "[A-Z]{2}"

test_strings %>%
  str_detect(pattern)
```

```{r}
pattern <- "[Cc]astle"

test_strings %>%
  str_detect(pattern, negate = TRUE)

# OR

pattern <- "Castle|castle"

test_strings %>%
  str_detect(pattern, negate = TRUE)

"
```

# Extracting matching substrings

[] - or operation
{1} - number of occurrences
(?i) - lets ignore capitalization
* - zero or more occurrence
+ - one or more occurrence

```{r}
strings <- c(
  "This string has an_address@email.com in it", 
  "This one has user.name@company.co.uk", 
  "Now we've got other_person_123@server.net and my.name@runningoutofideas.com"
)
```

```{r}
test_string <- "THIS is a string"

pattern <- "[(?i)a-z]"

test_string %>%
  str_extract(pattern)
```

```{r}
pattern <- "[a-z]"
strings %>%
  str_extract(pattern)

# ignored the capital letters at the start of the string
```

```{r}
pattern <- "[a-z]@"
strings %>%
  str_extract(pattern)

# extracts 
```

```{r}
invalid_email <- "gs@emailprovider.com"

pattern <- "[a-z]*@emailprovider.com"

invalid_email %>%
  str_extract(pattern)
```

```{r}
invalid_email <- "@emailprovider.com"
pattern <- "[a-z]+@emailprovider.com"

invalid_email %>%
  str_extract(pattern)

# Nothing infront of the @. Therefore nothing is returned 
```

```{r}
strings
```

```{r}
pattern <- "[a-z]+@"

strings %>%
  str_extract(pattern)
```

```{r}
pattern <- "[a-z0-9]+@[a-z]+"

strings %>%
  str_extract(pattern)
```

```{r}
pattern <- "[a-z0-9_\\.]+@[a-z\\.]+"

strings %>%
  str_extract(pattern)
```

```{r}
strings
```
```{r}
pattern <- "[a-z0-9_\\.]+@[a-z\\.]+"

extracted <- strings %>%
  str_extract_all(pattern)

extracted
```

```{r}
strings
```

```{r}
pattern <- "[a-z0-9_\\.]+@[a-z\\.]+"

strings %>%
  str_replace(pattern, "REDACTED")
```

```{r}
pattern <- "[a-z0-9_\\.]+@[a-z\\.]+"

strings %>%
  str_replace_all(pattern, "ooooweee")
```

# Task 1: 
Make a vector containing the strings:
“Colin’s mobile number definitely isn’t 07777777777.”
“A landline might be 01010 123456.”
“Mhairi, Del, Aileen and Steph have been some of your instructors so far.”
“I’m called your_name_here and I live in your_town_here.”
Use regular expressions to:
Extract all the names
Extract the landline number
Insert your name and where you live into the fourth string

```{r}
vector_strings <- c("Colin’s mobile number definitely isn’t 07777777777.", "A landline might be 01010 123456.", "Mhairi, Del, Aileen and Steph have been some of your instructors so far.", "I’m called your_name_here and I live in your_town_here.")

vector_strings
```

```{r}
pattern <- "[A-Z][a-z]+"

extracted <- vector_strings %>%
  str_extract_all(pattern)

extracted %>% flatten_chr()

# compresses the matrix, and gives us a list
```

```{r}
pattern <- "[0-9]{5} [0-9]{6}|[0-9]{11}"

vector_strings %>%
  str_extract_all(pattern) %>%
  flatten_chr()

# that extracts all the numbers there. Landline or Mobile
```

```{r}
my_name <- "Jonny"
my_address <- "Edinbro"

vector_strings %>%
  str_replace("your_name_here", my_name) %>%
  str_replace("your_town_here", my_address)
```

# Handy Regex Notations

# Standard Punctuation: ?!':;,

```{r}
punct_strings <- c(
  "Why is this a question?", 
  "I am a sentence with no punctation",
  "I am a well punctuated sentence.", 
  "WOW!"
)
punct_strings
```

```{r}
punct_strings %>%
  str_extract_all("[:punct:]")

# Gone into each vector and extracted the punctuation.
```

```{r}
punct_strings %>%
  str_extract_all("[A-Z]{3}[:punct:]") %>%
  flatten_chr()
```
#  Start of end of the string

^ - start of a string
$ - end of the string

```{r}
test_strings <- c(
  "abc", 
  "Abc",
  "abC",
  "s"
)

test_strings
```

```{r}
test_strings %>%
  str_extract_all("^[a-z]$") 

# Starts and ends with exactly one character of the e
```

```{r}
test_strings %>%
  str_extract_all("^[a-z]")

# First occurrence where is starts with a lowercase alphabet
```

```{r}
test_strings %>%
  str_extract_all("[a-z]$")

# First occurrence where is ends with a lowercase alphabet
```

```{r}
test_strings %>%
  str_extract_all("^[a-z]$")

# Only takes the observation that starts and ends with a letter of one character
```

```{r}
student <- c("Abigail", "Bjorn", "Caitlin")
age <- c(25, 56, 40)
emails <- c("abigal12@company.co.uk", "bsmith@gmail.co.uk", "caitlin_1980")

student_info <- tibble(student, age, emails)

student_info 

```

```{r}
student_info %>%
  mutate(email_detect = str_detect(emails, "[a-z0-9]+@"))
```

```{r}
student_info %>%
  mutate(email_new = str_extract(emails, "[a-z0-9_\\.]+@[a-z\\.]+"))
```

